Supervised Reinforcement Learning with Partial States: \\ a Teaching Method for Action Selection for HRI

### 2017-07-13 paper.tex
Update

\author{Emmanuel Senft \and S\'{e}verin Lemaignan\\
Centre for Robotics and Neural Systems \\
Plymouth University \\
United Kingdom\\
\And Paul Baxter\\
L-CAS\\
University of Lincoln\\
United Kingdom\\
 \And Tony Belpaeme\\
 CNRS\\ Plymouth University (UK) \\ iMinds \\ Ghent University (Be)}

@@

Human-Robot Interaction (HRI) aims at having humans and robot living together
in the society, interacting socially in many different environments and
contexts. Robot are expected to behave appropriately regardless of the domain
of interaction. However, not all behaviours are known in advance and can be
implemented in the robot before its deployment in the real world. Robots need
to be able to learn new tasks, new actions policies, new social norms and how
to make sense of the world surrounding them.

When interacting in the real world, humans face huge spaces of almost unlimited
dimensions, and despite this complexity, adults and even more children are able
to learn quickly, often from only a few demonstrations or examples. Humans are
able to take the most out of every datapoint availble to them. On the other
hand, recent advances in Machine Learning such as Deep Learning \cite{lecun2015deep}
have been made possible by the availability of huge dataset covering millions
of labeled examples or dozens of hours recording. These progress have
targetting either perception, where labeled datapoints can be available or
action selection in virtual environments where the impact of actions is
limited.

Similarly to humans, robots interacting in the real world will have to be able
to learn quickly, from different types of inputs: demonstrations, commands,
rewards and also autonomously.

State used for action selection can be from different sensors, continuous, discrete, representing a value or a state

\cite{kober2013reinforcement}
%===============================================================================

@@

The main framework to provide a robot with autonomous learning is Reinforcement
Learning (RL)(\cite{kober2013reinforcement,sutton1998reinforcement}. The main idea
behind reinforcement learning is that a agent is interacting in an environment
providing positive or numerical reward in reaction to the agent's actions and
the agent selects actions to maximise the expected cummulative discounted
reward.

Classic approaches RL relies on exploring virtual environments where the agent
tries to maximise the reward, but where in the end exploration often do not
have a real impact and where the agent can interact as long as needed to gather
enough information on the environment to find a correct action policy.

But in many real worl problem, relying only on random exploration to obtain
knowledge on the environment is not enough, the agent needs to have some initial
knowledge or additional knowledge provided often by humans to  work more
efficiently and achieve more complex results.

\subsection{Human enhanced Reinforcement Learning}
\subsubsection{Initial knowledge}
One way to use human knowledge to help an agent to learn is through providing
initial knowledge. 
alphago

IRL

OpenAI deep RL

One of the main challenge when applying RL to robotics is poor perfermance in
the early stages of the learning as the agent has to explore the environment to
gather initial knowledge. This challenge is especially
important for robotics as trials often have to happen in the real world (as
simulator close enough to the real world as not available) and as such can
present risks for the robot or for the persons interacting with it.

\subsubsection{Guiding the learning}
scafolding 
shaping
guidance

@@

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot as it would
have been passively or actvely validated by the supervisor. However this method
was not making use of any type of generalisation, it was directly mapping a
single state action pair to a reward.

@@

REFORMULATE: start with size of the world
To be usable in complex environments, a
learning algorithm requires generalisation as exactly identical spaces might
never been observed twice.

A classical approach to generalise in RL is to use a first layer of neural
network or to use deep reinforcement learning which relies on deep neural
networks to learn a mapping state-action-Qvalue for example. (ADDREF)
These approaches, neural network based, relies on having a large number of
datapoints to converge toward a good function approximator. However, as
discussed previously, in many application, these amounts of data are not
possible to be obtained and a robot might need to adapt to different persons and
be retrained quickly.

With that idea in mind, we propose to make a better use of human inputs, and ask
 the supervisor to identify which features of the environment are linked to the
 current selected action. That was, instead of having an state-action pair, we
 can have a partial-state action pair, with this partial-state being a subset of
 the state domain.

>>>>

\author{Emmanuel Senft \\
CNRS \\
Plymouth University \\
United Kingdom\\
\And S\'{e}verin Lemaignan\\
CNRS \\
Plymouth University \\
United Kingdom\\
\And Paul Baxter\\
L-CAS\\
University of Lincoln\\
United Kingdom\\
 \And Tony Belpaeme\\
 CNRS\\ Plymouth University (UK) \\ iMinds \\ Ghent University (Be)}

@@

Human-Robot Interaction (HRI) aims at having humans and robot living together
in the society, interacting socially in many different environments and
contexts. Robot are expected to behave appropriately regardless of the domain
of interaction. However, not all behaviours can be known in advance and can be
implemented in the robot before its deployment in the real world. Similarly to
 humans, robots need
to be able to learn new tasks, new actions policies, new social norms and how
to make sense of the world surrounding them.

When interacting in the real world, humans face huge spaces of almost unlimited
dimensions, and despite this complexity, adults and even more children are able
to learn quickly, often from only a few demonstrations or examples. Humans are
able to take the most out of every datapoint available to them. On the other
hand, recent impressive advances in Machine Learning such as Deep Learning~\cite{lecun2015deep}
have been made possible by the availability of huge dataset covering millions
of labeled examples or dozens of hours of recording. These progress have
targetting either perception, where labeled datapoints can be available or
action selection in virtual environments where the impact of actions is
limited.

Similarly to humans, robots interacting in the real world will have to be able
to learn quickly. One way of achieving this fast learning is to learn from
humans, i.e using human inputs to boostrap the learning. We argue that to be
capable to interact in the real world with humans, robots have to be able to
learn from different types of inputs: demonstrations, commands, rewards and 
also autonomously.

@@

The main framework to provide a robot with autonomous learning is Reinforcement
Learning (RL)~\cite{kober2013reinforcement,sutton1998reinforcement}. In RL, an
agent is interacting in an environment
providing positive or numerical reward in reaction to the agent's actions and
the agent learns an action policy to maximise the expected cummulative discounted
reward obtained by the agent.

Classic approaches RL relies on exploring virtual environments where the agent
tries to maximise the reward. In most of the cases where RL achieves success, it
had access to a virtual environment where the only real cost of exploring was
time and energy, and the agent can interact as long as needed to gather
enough information on the environment to find a correct action policy. When
these conditions are reunited, RL can achieve impressive results, as shown by
the victory of a computer in Backgammon back in the 90s~\cite{tesauro1995temporal} or more recently the success achieved in playing
Atari games based only on the image and the game rewards~\cite{mnih2015human}.

\subsection{Human enhanced Reinforcement Learning}

No simulator of world, and especially of social interactions allows robots to 
learn in a virtual environment, so robots, especially if the have to interact
with humans have to be able to learn in the real world.
However, in the real physical world, every actions have an impact on the
environment, especially in HRI when actions can upset and negatively impact
humans interacting with the robot. Additionally acquiring datapoints in can be
expensive in term of time and other ressources. As such many approaches aiming
at providing robot with learn to interact in the real world have relied on
knowledge provided by a human in different way to avoid this random exploration
presenting risk to the robot and persons interacting with it.

\subsubsection{Initial knowledge}
One way to use human knowledge to help an agent to learn is through providing
initial knowledge. By providing demonstrations of an expected actions policy,
humans can boostrap the learning and provide the agent with a more appropriate
action policy that it can improve over time.

These demonstrations can be used to create an initial action policy which could
have been discovered autonomously, but without requiring an initial discovery
phase where the performance has to be low when the agent is learning from its
environment. Recent work~\cite{hester2017learning} have shown that using initial
human demonstrations can allow the initial performance to be much better than
starting from scratch in playing Atari games, and even sometimes this allow to
achieve higher performance than only relying on exploration to learn an action
policy.

In some cases, the state space is too big to rely only on random exploration to
reach a successful action policy. For example the game of Go in its larger board
size contains more than $10^{210}$ states, so exhaustive search is not possible
with today's ressources. Authors of \cite{silver2016mastering} started with
supervised learning from go master's games to learn a decent policy and then
used convolutional networks, self play and tree search to achieve super-human
capabilities and beatting humans best players.

Another way to use these demonstrations is Inverse
Reinforcement Learning, the agent derives from demonstrations of an expert an
expected reward function that it can use to improve a policy based on the
demonstrations and so achieving higher performance than the demonstrations
themselves. By using this technic, Abbeel managed to teach an autonomous
helicopter to do high level acrobatics often only executable by high level
expert~\cite{abbeel2004apprenticeship}.

%One of the main challenge when applying RL to robotics is poor perfermance in
%the early stages of the learning as the agent has to explore the environment to
%gather initial knowledge. This challenge is especially
%important for robotics as trials often have to happen in the real world (as
%simulator close enough to the real world as not available) and as such can
%present risks for the robot or for the persons interacting with it.

\subsubsection{Guiding the learning}
Rather than providing initial knowledge to the agents, humans can also guide the
agent throughout the learning by organising the tasks the robot will face, this
approach is known as scaffolding \cite{saunders2006teaching}. This allows to
teach the robot successive more complex policies by increasing the complexity of
the task presented and reusing previously learnt behaviours.

shaping
guidance

@@

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot as it would
have been passively or actvely validated by the supervisor. However this method
was not making use of any type of generalisation, it was directly mapping a
single state action pair to a reward, and as such has limited efficiency in
environments with a continuous space or a high dimensional and undeterministic
one.

@@

REFORMULATE: start with size of the world and oragnisation by sensory inputs
To be usable in complex environments, a
learning algorithm requires generalisation as exactly identical spaces might
never been observed twice in the life time of the agent.

A classical approach to generalise in RL is to use a first layer of neural
network or to use deep reinforcement learning which relies on deep neural
networks to learn a mapping state action Qvalue for example. (ADDREF)
These approaches, neural network based, relies on having a large number of
datapoints to converge toward a good function approximator. However, as
discussed previously, in many application, these amounts of data are not
possible to be obtained and a robot might need to adapt to different persons and
be retrained quickly.

With that idea in mind, we propose to make a better use of human inputs, and ask
the supervisor to identify which features of the environment are linked to the
current selected action. Instead of having an state-action pair, we
can have a partial-state action pair, with this partial-state being a subset of
the state domain.
